Air cargo load and route planning in pickup and delivery operations

### 2023-06-10 CAIE/CIE.tex
Paper finished.

In aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading, for fast take-off, and mission accomplishment, especially in times of crisis, disaster relief missions, short deadlines for contracting customers, or any external pressure for immediate take-off. In addition, there are no commercially available systems that can assist the load and trip planners with the pallet building and aircraft loading and balancing demands for transport at each hub. This enables other risks, such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We defined and solved the problem of planning the loading and routing of a single aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs. This NP-hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs. We performed multiple experiments with a commercial solver and four well-known meta-heuristics on data based on the transport history of the {\it Brazilian Air Force}. We also designed a new heuristic that quickly finds good solutions for a wide range of problem sizes: an essential contribution, which solved all test scenarios within an acceptable operational run time.

@@

\subsubsection{With Shims heuristic}

@@

			\State $slack_i \gets V_i - vol_i$  \label{shims:beginsecond}
			\State $\eta_2 \gets \eta_1$  \label{shims:eta2a}
			\While{($\eta_2 \leq n_{\pi_k}$) {\bf and} ($ vol_i < (1 + 3 \times (1-limit)) \times V_i)$)}

@@

			\State $vol \gets 0$, $b \gets 1$, $shims_b \gets \varnothing$, $Set \gets \{ shims_b \}$  \label{shims:beginthird}

@@

				\If{$NewShims$} \label{new_shims2}
					\State $vol \gets 0$
					\State $b \gets b + 1$
					\State $shims_b \gets \{\ \}$
					\State $shims_b \gets shims_b \cup \{e_{ij}\}$
					\State $Set \gets Set \cup \{\ shims_b\ \}$
				\EndIf

@@

			\State $sh_w \gets shims$, where $shims \in Set$ and $\sum_{e_{ab} \in shims} w_b$ is maximum  \label{best_weight}		
			\State $sh_v \gets shims$, where $shims \in Set$ and $\sum_{e_{ab} \in shims} v_b$ is maximum \label{best_volume}
			\State $sh_{best} \gets shims$, where $shims \in \{sh_w, sh_v\}$ and $\sum_{e_{ab} \in x} s_b$ is maximum \label{best_score}

			\State $E_{N_{\pi_k}} \gets E_{N_{\pi_k}} \cup sh_{best}$ 	\label{shims:endthird}   \Comment{End of the third phase}

@@

	\item In {\it greedy phase}\/ (lines \ref{shims:pallets}-\ref{shims:endfirst}), a partial solution for each pallet $i$\/ is constructed by edges inclusion, until its volume $V_i \times limit$\/ is reached. The value $limit=0.92$\/ (line \ref{shims:limit}) was empirically determined by the {\it iRace} tool \cite{LopezIbanezManuel2016}. From lines \ref{shims:limit} to \ref{shims:endfirst}, as items are inserted, the value of $\eta_1$ is incremented and stops when the pallet volume reaches 92\% of its capacity. When it stops, this is the initial value of $\eta_2$, line \ref{shims:eta2a}.
	$\eta_2$, then it is incremented until the volume exceeds the pallet capacity with a factor of 1 + ( 2 x [1 - limit]).
	This factor of was obtained empirically, through many test with different problem instances. Higher values can improve the solution, but at the expense of performance. So we tried to keep it as low as possible, a trade-off between quality and performance.

	\item In {\it composition phase}\/ (lines \ref{shims:beginsecond}-\ref{shims:endsecond}), a set of shims named $Set$\/ is created for each pallet $i$, where each shim is formed by a set of edges in the range $[\eta_1,\eta_2]$, whose total volume is limited by $slack_i$. After the $\eta_2$ index, there may be items with little probability of being in the optimal solution because their low value of $\theta^{\pi_k}_{ij}$, being not considered for inclusion. In line \ref{shims:beginthird} the initial $Shim$ and $Set$ are created. In this phase, the heuristic that provided the best results, both in terms of time and quality, is based on {\it First-Fit Decreasing}, which is an approximation algorithm for the {\it Bin Packing Problem}\/ \cite{JohnsonGarey1985}. Basically, shims are created by accumulating the following edges, taking $slack_i$\/ as a limit.

@@

\begin{table}[H]
	\centering
	\caption{Items weight distribution ($kg$) in Brazil}  \label{tab:weights}
	\begin{tabular}{c c c c }
		\toprule
		$Group$ & $P$ & $low$ & $high$  \\		
		\midrule
		1              & 0.23           & 10  & 20  \\
		2              & 0.22           & 21  & 40  \\
		3              & 0.24           & 41  & 80  \\		
		4              & 0.23           & 81  & 200 \\
		5              & 0.08           & 201 & 340 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

@@

$ItemsGeneration$, which generates $N$, is described in Algorithm \ref{alg:itemsgen}. The parameter $scenario$\/ defines $L$\/ and $M$\/ (line \ref{ig:LM}), and the parameter $surplus$\/ sets a limit on the total volume of items at each node (line \ref{ig:extended}). To avoid simply loading all items, we use $surplus \in \{1.2, 1.5, 2.0\}$. This also represents more instances for tests in each scenario.

@@

{\color{blue}

@@

In the tests performed, we used a 64-bit, 16GB, 3.6GHz, eight-core processor with {\it Linux Ubuntu 22.04.1 LTS 64-bit} as the operational system and {\it Python 3.10.4}\/ as the programming language. We also used the well-known solver {\it Gurobi}\/ (www.gurobi.com), version 9.5.2.

We ran Algorithm \ref{alg:main} in the 5 scenarios described in Table \ref{tab:scenarios}, considering 6 methods for node-by-node solution: Gurobi (see \ref{solver}), ACO, NMO, TS, GRASP, and {\it Shims} (Algorithm \ref{alg:shims}). In generating the items in each node, we consider 3 values for the parameter $surplus$. 3 values for the time limit (1200s, 2400s, and 3600s).

The results obtained for the function $f$, with the corresponding run time in seconds, are shown in Tables \ref{tab:20} ($surplus = 1.2$), \ref{tab:50} ($surplus = 1.5$), \ref{tab:100} ($surplus = 2.0$), and in Table \ref{tab:100time} (different time limit comparisons).

For each $method$, $scenario$\/ and $surplus$, 7 different instances were generated. Therefore, 315 tests (5 scenarios $\times$\/ 3 volumes $\times$\/ 3 time limits $\times$\/ 7 instances) were performed for each method.

The average values were presented for $f_{\pi}$\/ and, for the {\bf run time (s)}, the average result obtained. To facilitate the comparison between the methods, we added a last column in these tables, where two values are indicated:
\begin{itemize}
	\item {\bf Normalized}: value between 0 and 1, which corresponds to the ratio between the sum of $f$\/ values obtained by the method in all scenarios and the sum of the best values obtained among all methods in all scenarios. The higher the value of {\bf Normalized}, the closer the method approached the best solutions found.
	\item {\bf Speed-up}: ratio of the sums of the run times of all scenarios and the sum of the method run times in all scenarios. The method with the highest {\bf Speed-up}\/ is the fastest.
\end{itemize}

	As the MIP solver was not capable of solving all scenarios in polynomial time or reached the computer's Read-Access Memory (RAM, 16 GiB) too early, we decided to set the Gurobi {\it MIPgap} parameter to 1\% to control the minimal quality of the returned solutions and abbreviate the solution time.
	From now on, we will refer to this Gurobi-specific model as {\it Gurobi* }.
	For more details on the Gurobi {\it MIPgap}, refer to \cite{mipgap}.

	This decision hinders the MIP solver from reaching optimal solutions in favor of the possibility of result comparisons with the {\it Shims} heuristic.

	The parameter $\pi_{TSP_1}$ represents the least cost tour (through an optimal TSP solution), and $\pi_{TSP_2}$, the opposite direction of travel, beginning and ending in the base. The values of $f$ are the best in these to solutions.

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\it surplus} = {\bf 1.2} and 3600s as time limit.}  \label{tab:20}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\bf Tours}&$method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ \specialcell{$\pi_{TSP_1}$\\ $\pi_{TSP_2}$ }} &\multirow{2}{*}{\it Gurobi* } & $f$ & 8.52    & 11.70   & 13.07   & 13.54 & 12.44     &  1.00\\ % 5927
		&                                                               &               {\bf run time (s)}  & 17      & 17      & 16      & 17    & 26        &  1.0 \\ % 93
		\cmidrule(lr){2-9}	
		&                                                               \multirow{2}{*}{\it Shims}    & $f$ & 8.54    & 11.63   & 12.97   & 13.43 & 12.36     &  0.99 \\ % 5893
		&                                                               &         {\bf run time (s)}        & 1       & 1       & 2       & 2     & 2         &  11.6 \\ % 8

@@

		\multirow{4}{*}{\it $S_{K}$ }&                                  \multirow{2}{*}{\it Gurobi* } & $f$ & 8.52    & 12.25   & 13.32   & 14.61  & x        & 1.00 \\ % 4869
		&                                                               &               {\bf run time (s)}  & 25      & 36      & 121     & 297    & x        &  1.0 \\ % 479
		\cmidrule(lr){2-9}	
		&                                                               \multirow{2}{*}{\it Shims}    & $f$ & 8.54    & 12.26   & 13.23   & 14.53  & 13.42    & 0.99  \\ % 4813
		&                                                               &         {\bf run time (s)}        & 1       & 2       & 6       & 17     & 106      & 18.4   \\ % 132
		\bottomrule
		\multicolumn{9}{c}{*{\it MIPgap} parameter set to 1\%. The $f$ values are \textpm 1\% distant from the optimal solution}\\

@@

{\color{blue}

	97\% of 16GiB was the RAM state before we aborted the {\it Gurobi*} execution on solving scenario 5 in Table \ref{tab:20}.

}

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\it surplus} = {\bf 1.5} and 3600s as time limit.}  \label{tab:50}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\bf Tours} & $method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ \specialcell{$\pi_{TSP_1}$\\ $\pi_{TSP_2}$ }} & \multirow{2}{*}{\it Gurobi* } & $f$ & 11.80 & 16.74 & 18.04 & 18.86 & 16.92 & 1.00 \\ % 8236
		&&               {\bf run time (s)}                                                                  & 38    & 30    & 27    & 29    & 81    & 1.00  \\ % 205
		\cmidrule(lr){2-9}	
		&\multirow{2}{*}{\it Shims}                                                                    & $f$ & 11.83 & 16.74 & 18.01 & 18.89 & 16.91 & 1.00  \\ % 
		&&         {\bf run time (s)}                                                                        & 1     & 2     & 2     & 3     & 3     & 20.5  \\ %

@@

		\multirow{4}{*}{\it $S_{K}$ } & \multirow{2}{*}{\it Gurobi* }                                  & $f$ & 11.81 & 17.01 & 18.25 & 20.59 & 18.36   & 1.00 \\ % 8602
		&&               {\bf run time (s)}                                                                  & 53    & 61    & 193   & 470   & 2241    & 1.0  \\ % 3018
		\cmidrule(lr){2-9}	
		&\multirow{2}{*}{\it Shims}                                                                    & $f$ & 11.83 & 17.00 & 18.21 & 20.45 & 18.00   & 0.99  \\ % 8549
		&&         {\bf run time (s)}                                                                        & 1     & 2     & 9     & 26    & 157     & 15.5 \\ % 195
		\bottomrule

@@

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\it surplus} = {\bf 2.0} and 3600s as time limit.}  \label{tab:100}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\bf Tours} & $method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ \specialcell{$\pi_{TSP_1}$\\ $\pi_{TSP_2}$ }} & \multirow{2}{*}{\it Gurobi* } & $f$ & 17.75 & 24.39 & 26.46 & 27.09 & 24.00 & 1.00 \\ % 11969
		&&                                                                               {\bf run time (s)}  & 162   & 95    & 74    & 68    & 67    & 1.0  \\ % 466
		\cmidrule(lr){2-9}	
		&                                                                   \multirow{2}{*}{\it Shims} & $f$ & 17.78 & 24.40 & 26.36 & 27.01 & 23.94 & 0.99  \\ % 11949
		&&                                                                                {\bf run time (s)} & 2     & 3     & 3     & 4     & 5     & 27.4  \\ % 17

@@

		\multirow{4}{*}{\it $S_{K}$ } & \multirow{2}{*}{\it Gurobi* } & $f$ & 17.75 & 25.24 & 26.46 & 29.11 &  x   & 1.00 \\ % 9856
		&&                                                                                {\bf run time (s)} & 151   & 123   & 319   & 689   &  x   & 1.0  \\ % 1282
		\cmidrule(lr){2-9}	
		&                                                                \multirow{2}{*}{\it Shims}    & $f$ & 17.78 & 25.25 & 26.38 & 28.98 & 26.08 & 1.00  \\ % 9839
		&&                                                                                {\bf run time (s)} & 2     & 3     & 13    & 37    & 229   & 23.3  \\ % 284
		\bottomrule

@@

{\color{blue}
	{\it Gurobi*} could not solve scenario 5 in Table \ref{tab:100} because $3600/K! = 0.71s$ is not enough time for {\it Gurobi*} to find a feasible node solution.
}

\subsubsection{Time limits comparisons}

\begin{table}[H]
	\centering
	\caption{Solutions with different time limitations for $S_K$ tours.}  \label{tab:100time_2}
	\footnotesize
	\begin{tabular}{ccc|ccc|ccc|ccc}
		\toprule
		&             \multicolumn{2}{r}{\bf Volume $surplus$} &\multicolumn{3}{c}{$1.2$}&\multicolumn{3}{c}{$1.5$}&\multicolumn{3}{c}{$2.0$} \\
		\midrule
		$method$& {\bf limit (s)} & {\bf Scenarios}&{\bf 3}&{\bf 4}&{\bf 5}&{\bf 3}&{\bf 4}&{\bf 5}&{\bf 3}&{\bf 4}&{\bf 5} \\
		\toprule

		\multirow{8}{*}{\it Gurobi } & \multirow{2}{*}{240*} & $f$ & 13.34 &  x      &  x      & 18.31 &  x      &  x      & x     & x       & x \\ % 
&               &  {\bf run time (s)}                              & 123   &  x      &  x      & 181   &  x      &  x      & x     & x       & x \\ % 
\cmidrule(lr){2-12}

		&                              \multirow{2}{*}{1200} & $f$ & 13.31 & 14.67   & x       & 18.25 &  20.64  &  x      & 26.49 & 27.97   & x \\ % 
		&        &  {\bf run time (s)}                             & 129   & 217     & x       & 190   &  304    &  x      & 384   & 579     & x \\ % 
		\cmidrule(lr){2-12}                              
		& \multirow{2}{*}{2400}& $f$                               & 13.30 & 14.59   & 13.59   & 18.25 &  20.60  &  x      & 26.46 & 29.10   & x \\ % 
		&        &  {\bf run time (s)}                             & 127   & 313     & 1444    & 188   &  327    &  x      & 381   & 543     & x \\ % 
		\cmidrule(lr){2-12}

		& \multirow{2}{*}{3600}& $f$                               & 13.32 & 14.61   & x       & 18.25 &  20.59  & 18.36   & 26.46 & 29.11   & x \\ % 
		&        &  {\bf run time (s)}                             & 121   & 297     & x       & 193   &  470    & 2241    & 319   & 689     & x \\ %

>>>>

In aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading, for fast take-off, and mission accomplishment, especially in times of crisis, disaster relief missions, short deadlines for contracting customers, or any external pressure for immediate take-off. {\color{blue} In addition, there are no commercially available systems that can assist load and trip planners with pallet building and aircraft-balanced loading plans, with demands for transport at each hub.} This enables other risks, such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We defined and solved the problem of planning the loading and routing of a single aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs. This NP-hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modeled using standardized pallets in fixed positions, obeying center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs. We performed multiple experiments with a commercial solver and four well-known meta-heuristics on data based on the transport history of the {\it Brazilian Air Force}. {\color{blue}We also designed a new heuristic that quickly finds practical solutions for a wide range of problem sizes, an essential contribution that solved all test scenarios within a much shorter time than is operationally acceptable.}

@@

\subsubsection{With the Shims heuristic}

@@

			\State $slack_i \gets V_i - vol_i$  \label{shims:beginsecond}
			\State $\eta_2 \gets \eta_1$  \label{shims:eta2a}
			\While{($\eta_2 \leq n_{\pi_k}$) {\bf and} ($ vol_i < (1 + 2 \times (1-limit)) \times V_i$)}

@@

			\State $vol \gets 0$; $b \gets 1$; $shims_b \gets \varnothing$; $Set \gets \{ shims_b \}$  \label{shims:beginthird}

@@

				\If{$NewShims$} \label{new_shims2}
					\State $vol \gets 0$; $b \gets b + 1$;  $shims_b \gets \{e_{ij}\}$
					\State $Set \gets Set \cup \{ shims_b \}$
				\EndIf

@@

			\State $sh_w \gets shims$, where $shims \in Set$ and $\sum_{e_{ij} \in shims} w_j$\/ is maximum  \label{best_weight}		
			\State $sh_v \gets shims$, where $shims \in Set$ and $\sum_{e_{ij} \in shims} v_j$\/ is maximum \label{best_volume}
			\State $sh_{best} \gets shims$, where $shims \in \{sh_w, sh_v\}$\/ and $\sum_{e_{ij} \in shims} s_j$\/ is maximum \label{best_score}
			\State $E_{N_{\pi_k}} \gets E_{N_{\pi_k}} \cup sh_{best}$ \label{shims:endthird}  \Comment{End of the third phase}

@@

	\item In {\it greedy phase}\/ (lines \ref{shims:pallets}-\ref{shims:endfirst}), a partial solution for each pallet $i$\/ is constructed by adding edges following $\theta_{ij}$\/ order. The $\eta_1$\/ index corresponds to the accumulated volume equal to $V_i \times limit$. In $\eta_2$\/ index, this same accumulated volume reaches $(1 + 2\times(1 - limit))\times V_i$. The size of this range $[\eta_1,\eta_2]$\/ was defined empirically, and the value $limit=0.92$\/ (line \ref{shims:limit}) was determined by the {\it iRace} tool \cite{LopezIbanezManuel2016}.

	\item In {\it composition phase}\/ (lines \ref{shims:beginsecond}-\ref{shims:endsecond}), a set of shims named $Set$\/ is created for each pallet $i$, where each shim is formed by a set of edges in the range $[\eta_1,\eta_2]$, whose total volume is limited by $slack_i$. In this phase, the heuristic that provided the best results, both in terms of time and quality, is based on {\it First-Fit Decreasing}, which is an approximation algorithm for the {\it Bin Packing Problem}\/ \cite{JohnsonGarey1985}. Basically, shims are created by accumulating the following edges, taking $slack_i$\/ as a limit.

@@

\begin{table}[H]
	\centering
	\caption{Items weight distribution}  \label{tab:weights}
	\begin{tabular}{c c c c }
		\toprule
		$Group$ & $P$ & $low$ ($kg$) & $high$ ($kg$) \\		
		\midrule
		1              & 0.23           & 10  & 20  \\
		2              & 0.22           & 21  & 40  \\
		3              & 0.24           & 41  & 80  \\		
		4              & 0.23           & 81  & 200 \\
		5              & 0.08           & 201 & 340 \\
		\bottomrule
	\end{tabular}
\end{table}

@@

$ItemsGeneration$, which generates $N$, is described in Algorithm \ref{alg:itemsgen}.

@@

$scenario$\/ defines $L$\/ and $M$\/ (line \ref{ig:LM}), and $surplus$\/ sets a limit on the total volume of items at each node (line \ref{ig:extended}). To avoid simply loading all items, we use $surplus \in \{1.2, 1.5, 2.0\}$. This also represents more instances for tests in each scenario.

@@

@@

In the tests performed, we used a 64-bit, 16GB, 3.6GHz, eight-core processor with {\it Linux Ubuntu 22.04.1 LTS 64-bit} as the operational system and {\it Python 3.10.4}\/ as the programming language. We also used the well-known solver $Gurobi$\/ ({\tt www.gurobi.com}), version 9.5.2.

We ran Algorithm \ref{alg:main} considering 5 values for $scenario$\/ from Table \ref{tab:scenarios}, 3 values for  $surplus$\/ from $\{1.2, 1.5, 2.0\}$, 3 values for $time$\/ from $\{1200s, 2400s, 3600s\}$, and 6 different $method$\/ for the node-by-node solution: $Gurobi$\/ (see \ref{solver}), ACO, NMO, TS, GRASP, and $Shims$\/ (Algorithm \ref{alg:shims}).

In order for $Gurobi$\/ to be able to solve the largest possible number of tests without memory overflow, we set its parameter {\it MIPgap}\/ to 1\%. This shortens its runtime, in addition to ensuring that its objective function $f$\/ is at most 1\% of the optimal solution. For more details, see {\tt https://support.gurobi.com}.

Among the heuristics, we will present only the results obtained by $Shims$, which had the best performance. For each $scenario$, $surplus$\/ and $time$, 7 different instances were generated. We will present the average of the objective function $f$\/ and the runtime of $Gurobi$\/ and $Shims$\/ in these instances. To facilitate the comparison between both, we added a last column in the tables, where two values are indicated:
\begin{itemize}
	\item {\bf Normalized}: value between 0 and 1, which corresponds to the ratio between the sum of $f$\/ values obtained by the method in all scenarios and the sum of the best values obtained by both methods in all scenarios. The higher the value of {\bf Normalized}, the closer the method approached the best solutions found.
	\item {\bf Speed-up}: ratio of the sums of the runtimes of all scenarios and the sum of the method runtimes in all scenarios. The method with the highest {\bf Speed-up}\/ is the fastest.
\end{itemize}

We also indicate the two adopted strategies: dedicating all the processing time to the two shortest tours, or distributing it among all the $K!$\/ tours.

The results obtained with $time = 3600s$, which is the highest tested time limit, are in Tables \ref{tab:20}, \ref{tab:50} and \ref{tab:100}, with $surplus$\/ values $1.2$, $1.5$\/ and $2.0$, respectively. We indicate with an {\bf x} the cases where $Gurobi$\/ did not find a feasible solution within this time limit, or had to be aborted due to high RAM consumption.

\begin{table}[H]
	\centering
	\caption{Solutions with $surplus = 1.2$\/ and $time =  3600s$}  \label{tab:20}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\specialcell{{\bf Tested}\\{\bf tours}}}&$method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{Shortest} &\multirow{2}{*}{\it Gurobi } & $f$ & 8.52    & 11.70   & 13.07   & 13.54 & 12.44     &  1.00\\ % 5927
		&                                                               &               {time (s)}  & 17      & 17      & 16      & 17    & 26        &  1.0 \\ % 93
		\cmidrule(lr){2-9}	
		&                                                               \multirow{2}{*}{\it Shims}    & $f$ & 8.54    & 11.63   & 12.97   & 13.43 & 12.36     &  0.99 \\ % 5893
		&                                                               &         {time (s)}        & 1       & 1       & 2       & 2     & 2         &  11.6 \\ % 8

@@

		\multirow{4}{*}{All $K!$}&                                  \multirow{2}{*}{\it Gurobi } & $f$ & 8.52    & 12.25   & 13.32   & 14.61  & x        & 1.00 \\ % 4869
		&                                                               &               {time (s)}  & 25      & 36      & 121     & 297    & x        &  1.0 \\ % 479
		\cmidrule(lr){2-9}	
		&                                                               \multirow{2}{*}{\it Shims}    & $f$ & 8.54    & 12.26   & 13.23   & 14.53  & 13.42    & 0.99  \\ % 4813
		&                                                               &         {time (s)}        & 1       & 2       & 6       & 17     & 106      & 18.4   \\ % 132
		\bottomrule

@@

\begin{table}[H]
	\centering
	\caption{Solutions with $surplus = 1.5$\/ and $time =  3600s$}  \label{tab:50}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\specialcell{{\bf Tested}\\{\bf tours}}} & $method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ Shortest } & \multirow{2}{*}{\it Gurobi } & $f$ & 11.80 & 16.74 & 18.04 & 18.86 & 16.92 & 1.00 \\ % 8236
		&&               {time (s)}                                                                  & 38    & 30    & 27    & 29    & 81    & 1.00  \\ % 205
		\cmidrule(lr){2-9}	
		&\multirow{2}{*}{\it Shims}                                                                    & $f$ & 11.83 & 16.74 & 18.01 & 18.89 & 16.91 & 1.00  \\ % 
		&&         {time (s)}                                                                        & 1     & 2     & 2     & 3     & 3     & 20.5  \\ %

@@

		\multirow{4}{*}{All $K!$} & \multirow{2}{*}{\it Gurobi }                                  & $f$ & 11.81 & 17.01 & 18.25 & 20.59 & 18.36   & 1.00 \\ % 8602
		&&               {time (s)}                                                                  & 53    & 61    & 193   & 470   & 2241    & 1.0  \\ % 3018
		\cmidrule(lr){2-9}	
		&\multirow{2}{*}{\it Shims}                                                                    & $f$ & 11.83 & 17.00 & 18.21 & 20.45 & 18.00   & 0.99  \\ % 8549
		&&         {time (s)}                                                                        & 1     & 2     & 9     & 26    & 157     & 15.5 \\ % 195
		\bottomrule

@@

\begin{table}[H]
	\centering
	\caption{Solutions with $surplus = 2.0$\/ and $time =  3600s$}  \label{tab:100}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\specialcell{{\bf Tested}\\{\bf tours}}}& $method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ Shortest} & \multirow{2}{*}{\it Gurobi } & $f$ & 17.75 & 24.39 & 26.46 & 27.09 & 24.00 & 1.00 \\ % 11969
		&&                                                                               {time (s)}  & 162   & 95    & 74    & 68    & 67    & 1.0  \\ % 466
		\cmidrule(lr){2-9}	
		&                                                                   \multirow{2}{*}{\it Shims} & $f$ & 17.78 & 24.40 & 26.36 & 27.01 & 23.94 & 0.99  \\ % 11949
		&&                                                                                {time (s)} & 2     & 3     & 3     & 4     & 5     & 27.4  \\ % 17

@@

		\multirow{4}{*}{All $K!$} & \multirow{2}{*}{\it Gurobi } & $f$ & 17.75 & 25.24 & 26.46 & 29.11 &  x   & 1.00 \\ % 9856
		&&                                                                                {time (s)} & 151   & 123   & 319   & 689   &  x   & 1.0  \\ % 1282
		\cmidrule(lr){2-9}	
		&                                                                \multirow{2}{*}{\it Shims}    & $f$ & 17.78 & 25.25 & 26.38 & 28.98 & 26.08 & 1.00  \\ % 9839
		&&                                                                                {time (s)} & 2     & 3     & 13    & 37    & 229   & 23.3  \\ % 284
		\bottomrule

@@

From these data, we can make some conclusions:
\begin{itemize}
\item The strategy of testing all $K!$\/ often gives a better quality solution.
\item $Gurobi$\/ fails in some cases, when $scenario=5$\/ and the strategy is to check all $K!$\/ tours. This occurs because the time limit per node is smaller, and there tends to be more packed contents on the aircraft, reducing the space for allocating items and making the solution difficult.
\item When $Gurobi$\/ finishes, it finds the best solution, but the one obtained by $Shims$\/ reaches $99\%$\/ of that value.
\item $Shims$\/ always finds a solution, being 11 to 27 times faster.
\item All runtimes are well under the time limit because the solution on many nodes can be fast. Anyway, in all the tests performed, the maximum time spent by $Shims$\/ did not reach 4 minutes, that is, it is practically instantaneous. On the other hand, when $scenario=5$\/ and $surplus=1.5$, Gurobi spent almost 40 minutes.
\end{itemize}

Table \ref{tab:100time} shows the results obtained with the strategy of testing the $K!$\/ tours in scenarios 4 and 5, with different time limits. We can observe more cases where $Gurobi$\/ fails. When $Gurobi$ finishes, $Shims$\/ finds a solution of similar quality in much less time.

\begin{table}[H]
	\centering
	\caption{Solutions in scenarios 4 and 5 testing all $K!$\/ tours}  \label{tab:100time}
	\footnotesize
	\begin{tabular}{ccc|cc|cc|cc}
		\toprule
		&             \multicolumn{2}{c}{$surplus$} &\multicolumn{2}{c}{$1.2$}&\multicolumn{2}{c}{$1.5$}&\multicolumn{2}{c}{$2.0$} \\
		\midrule
		$method$& {\specialcell{$time$ (s)}} & {\bf Scenarios}    & {\bf 4} & {\bf 5} & {\bf 4} & {\bf 5} & {\bf 4} & {\bf 5} \\
		\toprule

		\multirow{6}{*}{\it Gurobi }& \multirow{2}{*}{1200} & $f$ & 14.67   & x       &  20.64  &  x      & 27.97   & x \\ % 
		&        &  {time (s)}                             & 217     & x       &  304    &  x      & 579     & x \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$                               & 14.59   & 13.59   &  20.60  &  x      & 29.10   & x \\ % 
		&        &  {time (s)}                             & 313     & 1444    &  327    &  x      & 543     & x \\ % 
		\cmidrule(lr){2-9}                              		
		& \multirow{2}{*}{3600}& $f$                               & 14.61   & x       &  20.59  & 18.36   & 29.11   & x \\ % 
		&        &  {time (s)}                             & 297     & x       &  470    & 2241    & 689     & x \\ % 		
		\midrule
		\multirow{6}{*}{\it Shims}   & \multirow{2}{*}{1200} & $f$ & 14.53   & 13.42   &  20.45  & 18.00   & 28.98   & 26.08 \\ % 
		&        & {time (s)}                              & 18      & 109     &  26     & 160     & 37      & 203   \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$                               & 14.53   & 13.42   &  20.45  & 18.00   & 28.98   & 26.08 \\ % 
		&        &  {time (s)}                             & 18      & 107     &  25     & 159     & 37      & 192   \\ % 	
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{3600} & $f$                              & 14.53   & 13.42   &  20.45  & 18.00   & 28.98   & 26.08 \\ %  
		&        &  {time (s)}                             & 17      & 106     &  26     & 157     & 37      & 181   \\

\begin{table}[H]
	\centering
	\caption{Solutions with different time limitations for $S_K$ tours.}  \label{tab:100time_2}
%\footnotesize
%	\tiny
\scriptsize	
	\setlength{\tabcolsep}{3.5pt}

	\begin{tabular}{ccc|ccccc|ccccc|ccccc}
		\toprule
		&             \multicolumn{2}{r}{\bf Volume $surplus$} &\multicolumn{5}{c}{$1.2$}&\multicolumn{5}{c}{$1.5$}&\multicolumn{5}{c}{$2.0$} \\
		\midrule
		$method$& \specialcell{\bf limit\\ (s)} & {\bf Scenarios}  &{\bf 1}&{\bf 2}&{\bf 3}&{\bf 4}&{\bf 5}&{\bf 1}&{\bf 2}&{\bf 3}&{\bf 4}&{\bf 5}&{\bf 1}&{\bf 2}&{\bf 3}&{\bf 4}&{\bf 5} \\
		\toprule

		\multirow{10}{*}{\it Gurobi } & \multirow{2}{*}{240*} & $f$ & 8.53 & 12.24 & 13.34 &  x      &  x      & 11.81 & 17.02 & 18.31 &  x      &  x      & 17.75 & 25.24 & x     & x       & x \\ % 
		&               &  {\bf time (s)}                              & 41   & 43    & 123   &  x      &  x      & 50    & 62    & 181   &  x      &  x      & 166   & 140   & x     & x       & x \\ % 
		\cmidrule(lr){2-18}

		&                              \multirow{2}{*}{1200} & $f$ & 8.52 & 12.25 & 13.31 & 14.67   & x       & 11.81 & 17.02 & 18.25 &  20.64  &  x      & 17.75 & 25.24 & 26.49 & 27.97   & x \\ % 
		&        &  {\bf time (s)}                             & 39   & 41    & 129   & 217     & x       & 52    & 61    & 190   &  304    &  x      & 161   & 139   & 384   & 579     & x \\ % 
		\cmidrule(lr){2-18}                              
		& \multirow{2}{*}{2400}& $f$                               & 8.53 & 12.25 & 13.30 & 14.59   & 13.59   & 11.80 & 17.02 & 18.25 &  20.60  &  x      & 17.75 & 25.24 & 26.46 & 29.10   & x \\ % 
		&        &  {\bf time (s)}                             & 33   & 39    & 127   & 313     & 1444    & 51    & 62    & 188   &  327    &  x      & 158   & 127   & 381   & 543     & x \\ % 
		\cmidrule(lr){2-18}

		& \multirow{2}{*}{3600}& $f$                               & 8.52 & 12.25 & 13.32 & 14.61   & x       & 11.81 & 17.01 & 18.25 &  20.59  & 18.36   & 17.75 & 25.24 & 26.46 & 29.11   & x \\ % 
		&        &  {\bf time (s)}                             & 25   & 36    & 121   & 297     & x       & 53    & 61    & 193   &  470    & 2241    & 151   & 123   & 319   & 689     & x \\ %
